---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r setup, include=FALSE}
require(data.table)
unzip(zipfile="activity.zip")
data <- as.data.table(read.csv("activity.csv"))
```

## What is mean total number of steps taken per day?

```{r setup, include=TRUE}
library(ggplot2)
sum.steps <- ddata[, sum(steps, na.rm = TRUE), by=date]
setnames(sum.steps, "V1", "Daily.steps")
hist(sum.steps[,Daily.steps], breaks = 50, xlab="Number of steps taken each day")
mean(sum.steps[,Daily.steps], na.rm=TRUE)
median(sum.steps[,Daily.steps], na.rm=TRUE)
```


## What is the average daily activity pattern?

```{r setup, include=TRUE}
avg.steps.interval  <- data[, mean(steps, na.rm=TRUE), by=interval]
setnames(avg.steps.interval, "V1", "Interval.steps")
plot(avg.steps.interval[,interval] , avg.steps.interval[,Interval.steps], type='l', ylab="Average number of steps", xlab = "5-minute intervall" )
avg.steps.interval[max(Interval.steps),interval]
```


## Imputing missing values

Total number of NA's in the dataset

```{r setup, include=TRUE}
nrow(data[is.na(steps),])
```

Replace the NA's with the average number of steps in that 5-minute intervall over all days. 

```{r setup, include=TRUE}
new.data <- merge(data, avg.steps.interval, by="interval", all.x=TRUE) 
new.data[,steps:=as.numeric(steps)]
new.data[is.na(steps), steps := Interval.steps]
new.data[,Interval.steps:=NULL]
```
Histogram, mean and median with data with NA's replaced 

```{r setup, include=TRUE}
new.data <- merge(data, avg.steps.interval, by="interval", all.x=TRUE) 
new.data[,steps:=as.numeric(steps)]
new.data[is.na(steps), steps := Interval.steps]
new.data[,Interval.steps:=NULL]
```

After replacing the NA values with mean number of steps in a 5-minute interval the mean and median increase.

## Are there differences in activity patterns between weekdays and weekends?

Adding two factos to the data. One factor indicating if the day is a weekday and one idicating a weekend.

```{r setup, include=TRUE}
data[, day := weekdays(as.Date(date))]
data[, weekday := (day=="Monday"|day=="Tuesday"|day=="Wednesday"|day=="Thursday"|day=="Friday")*1]
data[, weekend := (day=="Saturday"|day=="Sunday")*1]
```

Panel plot containing a time series plot of the 5-minute interval and the average number of steps taken.


```{r setup, include=TRUE}
par(mfcol=c(2,1))
avg.steps.interval.weekday  <- data[weekday==1, mean(steps, na.rm=TRUE), by=interval]
setnames(avg.steps.interval.weekday, "V1", "Interval.steps")
plot(avg.steps.interval.weekday[,interval] , avg.steps.interval.weekday[,Interval.steps], type='l', ylab="Average number of steps", xlab = "5-minute intervall" , main = "Average Weekday")
avg.steps.interval.weekday[max(Interval.steps),interval]

avg.steps.interval.weekend  <- data[weekend==1, mean(steps, na.rm=TRUE), by=interval]
setnames(avg.steps.interval.weekend, "V1", "Interval.steps")
plot(avg.steps.interval.weekend[,interval] , avg.steps.interval.weekend[,Interval.steps], type='l', ylab="Average number of steps", xlab = "5-minute intervall" , main = "Average Weekend")
avg.steps.interval.weekend[max(Interval.steps),interval]
```
